Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-08-28 Manuscript/manuscript.tex
MS: tiny change

In this work, we empirically sample this parameter space $f_q \in [0.75, 1.0]$
to find an atomistic MD simulation force field parameters for lipids that
accurately describe the lipid head group response to Na$^+$ and Ca$^{2+}$ concentrations
when compared with NMR data \cite{akutsu81,altenbach84}. These data can be used to accurately asses
ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
We chose Lipid14~\cite{dickson14} force field parameters as a starting point,
because their response to bound ions was apparently most realistic against NMR data
in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
Also glycerol backbone and head group structures in Lipid14 model were
relatively realistic when compared with other state of the art lipid models \cite{botan15}.
The ECC correction was applied to Lipid14 parameters by scaling partial charges of the headgroup,
glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
are expected to have the largest contribution to the cation binding.
The hydrocarbon chain parameters are not modified, because they are
already highly optimized and give generally a good description for
hydrophobic part of lipid bilayers in various conditions
in most lipids models, including Lipid14 \cite{ollila16}.
In contrast, improvements in glycerol backbone and headgroup
parameters are required in all available lipid models~\cite{botan15}.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities2.eps}
  \caption{\label{fig:cacl-dens}
    Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}.
    Relative surface excess for the respective simulations are: 
    Lipid14 with \AA{}qvist ions $\Gamma_i^w = XX nm^{-3}$; 
    Lipid14 with Dang ions $\Gamma_i^w = 0.3 nm^{-3}$; 
    ECC-lipids with ECC-ions $\Gamma_i^w = 0.07 nm^{-3}$. 
    %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
    Densities of~\ce{Cl^-} and water are divided with 2 and 200, respectively, to visalize them
    with the same scale as \ce{Ca^{2+}}. The molar concentration of the ions in water is 350~mM in all systems
    presented here. 
    }
  \todo{Calculate the surface excess for L14+Aqvist sim.}
  \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
  \todo{JOE: Change the figure so that it contains a membrane background} \\
  \todo{Resuls with Lipid14+Dang to be added once done -- running with larger box.}
\end{figure}

@@

In addition, we estimated the relative binding affinities of several moieties in POPC:
phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
Using the probability isodensity contours (see Fig.~\ref{fig:volmaps}), 
we estimate that the largest contribution
to the binding of \ce{Ca2+} to POPC membranes likely comes from the phosphate group. 
Although the isodensity plots are relatively easy to interpret,
the contours shown in Fig.~\ref{fig:volmaps} cannot conclusively tell on the details
of \ce{Ca2+} binding to any of the two carbonyl moieties.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=6.0cm]{../Fig/volmap_resid10_Ca_Cl_PO4Cent.png} \\
  \includegraphics[width=4.0cm]{../Fig/volmap_resid10_Ca_Cl_sn1Cent.png}
  \includegraphics[width=3.6cm]{../Fig/volmap_resid10_Ca_Cl_sn2Cent.png}
  \caption{\label{fig:volmaps}
      Contours of probability isodensities of \ce{Ca2+} with respect to 
      various moieties fixed in space (highlighted with transparent spheres): 
      phosphate moiety, side chain 1 carbonyl group and side chain 2 carbonyl group.
      Shown contours suggest that the dominant contribution 
      to \ce{Ca2+} binding likely comes from the phosphate group, 
      but remain inconclusive on the details of the interactions with any of the two carbonyl groups. 
  }
\end{figure}

>>>>

In this work, we empirically sample this parameter space $f_q \in [0.75, 1.0]$
to find an atomistic MD simulation force field parameters for lipids that
accurately describe the lipid head group response to varying concentrations of cations
when compared to NMR data \cite{akutsu81,altenbach84,scherer89}. Such data can be used to accurately asses
ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
We chose Lipid14~\cite{dickson14} force field parameters as a starting point,
because their response to bound ions was apparently most realistic against NMR data
in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
Also glycerol backbone and head group structures in Lipid14 model were
relatively realistic when compared with other state of the art lipid models \cite{botan15}.
The ECC correction was applied to Lipid14 parameters by scaling partial charges of the headgroup,
glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
are expected to have the largest contribution to the cation binding.
The hydrocarbon chain parameters are not modified, because they are
already highly optimized and give generally a good description for
hydrophobic part of lipid bilayers in various conditions
in most lipids models, including Lipid14 \cite{ollila16}.
In contrast, improvements in glycerol backbone and headgroup
parameters are required in all available lipid models~\cite{botan15}.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=9.0cm,angle=0]{../Fig/CAdensities2.eps}
  \caption{\label{fig:cacl-dens}
    Number density of \ce{Ca^{2+}} and \ce{Cl^-} as a function of membrane normal axis
    for different force fields. Data for Lipid14 with \AA{}qvist ions are taken directly from Ref. \citenum{catte16}
    %Normalization factor for ions is 1 for monovalent ions (i.e.~\ce{Na^+} and~\ce{Cl^-}),
    Densities of~\ce{Cl^-}) and water are divided with 2 and 200, respectively, to visalize them
    with the same scale as \ce{Ca^{2+}}. The molar concentration in water is 350~mM in all systems
    presented here. 
    }
  \todo{PAVEL: draw phosphate position with its variance, add water density (scaled) and include the number of$\Gamma$-surface access.} \\
  \todo{JOE: Change the figure so that it contains a membrane background} \\
  \todo{Resuls with Lipid14+Dang to be added once done.}
\end{figure}

@@

@@
